\singlespacing \vspace{-1.6cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2022-01-31 Manuscript/Lanuza_et_al_2021_Ecol_Letters.Rmd
Edits to fit the max of 5000 words

\textbf{Number of words:} Abstract  (146 words), main text (5097 words).

\textbf{Number of references:} 77 (CHECK! Probably less now)

@@

Plant life-history strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of plant trait variation, little is known about plant reproductive trade-offs and how these constrain life-history strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these drive interactions with floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over half of all plant reproductive trait variation was explained by two independent axes. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. In addition, plant reproductive trade-offs determined important differences in the interaction level among floral visitor guilds. Our study shows the main plant reproductive trade-offs and their relevance to understand plant-pollinator interactions in a global context.

@@

Despite the astonishing diversity of floral structures among flowering plants [@barrett2002; @schiestl2013] and their importance in shaping plant-pollinator interactions [@fenster2004; @dellinger2020], a unified framework that **describes the major** plant reproductive trade-offs is currently lacking [@roddy2021]. In addition, macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with floral visitors at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. **However, plant reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020] and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations [@tur2013; @devaux2014].**

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases [e.g., TRY, @kattge2011; and COMPADRE, @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies have focused on vegetative traits such as leaf [@wright2004], wood [@chave2009], or root [@laughlin2021] trade-offs with little or no attention given to reproductive traits [@evojtko2020; @roddy2021] which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, [@barrett2003; @moeller2017] and outcrossing levels are positively correlated with flower size [@goodwillie2010]. In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species’ reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators [@ollerton2011; @rodger2021]. However, it is still unknown to what extent these different reproductive compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties [@bartomeus2013; @olito2015; @rowe2020]. Moreover, plant traits can define species’ network roles [e.g., specialists vs generalists\; @lazaro2013; @tur2013]. For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. Remarkably, the combination of traits has shown to increase the predictive power of the network interactions [@eklof2013]. **Moreover, the position in the trait space can determine plant and pollinator species’ functional roles [@dehling2016]. Therefore, by considering the multidimensional reproductive trait space we could progress our understanding of plant-pollinator associations. Importantly, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.**

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (https://www.ncbi.nlm.nih.gov/taxonomy) for plants and ITIS (https://www.itis.gov/) for pollinators, using the R package _taxize_ [@chamberlain2020] version _0.9.99_. We filled the ‘not found’ searches manually using http://www.theplantlist.org/ and http://www.mobot.org/ for plants and http://www.catalogueoflife.org/ for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. Note that autonomous selfing was recorded quantitatively (fruit set) and qualitatively because this trait was rarely available in its numerical form. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns x 1506 species) we were able to fill 24,341 cells (80.8% of the dataset, see Fig. S1 for missing values information for each trait).

@@

n_of_records <- subset(missing.values, isna=="FALSE")
Trait <- linebreak(c("Plant height \n (m)", "Flower width \n (mm)", "Flower length \n (mm)", "Inflorescence width \n (mm)", "Style length \n (mm)", "Ovules per flower",
           "Flowers per plant", "Nectar \n ($\\mu$l)", "Nectar \n (mg)", "Nectar concentration \n ($\\%$)", "Pollen grains \n per flower", "Autonomous selfing \n (fruit set)"),align = c("c"))

Trait1 <- linebreak(c("Lifepan", "Life form", "Flower \n shape", "Flower \n symmetry", "Nectar", "Compatibility \n system", "Breeding \n system", "Autonomous \n selfing", " ", " ", " ", " "),align = c("c"))

T.cat <- linebreak(c("Short-lived \n Perennial", "Herb \n Shrub \n Tree", "Brush \n Campanulate \n Capitulum \n Open \n Papilionaceous \n Tube",  "Actinomorphic \n Zygomorphic",
           "Presence \n Absence", "Self-incompatible \n Part. self-compatible \n Self-compatible", "Hermaphrodite \n Monoecious \n Dioecious","None \n Low \n Medium \n High", " ", " ", " ", ""),align = c("l"))

@@

Type_1 <- linebreak(c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Reproductive \n biology", "Reproductive \n biology","Reproductive \n biology","","","", ""),align = c("c"))

@@

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE, align = c("c","l","c","c","l","l","c"),caption = "\\textbf{Table 1} Plant functional traits used in this study.") %>%
kable_styling(full_width = F,font_size = 10) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 3),bold=T)  %>%
    column_spec(c(1,3), bold = T) %>% row_spec(0, bold=T)

@@

Trait missing values were imputed with the function _missForest_ [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of _missForest_ [@penone2014]. To extract the eigenvectors, we used the function _PVRdecomp_ from the package _PVR_ [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. However, we were unable to include nectar and pollen traits on the imputation process because of the high percentage of missing values (Fig. S1). Hence, the imputed dataset had 1,506 species, seven categorical and eight numerical variables and 5.79% of missing values. Further, we conducted an additional imputation process on the subset of species with data for pollen per flower and microliters of nectar. This subset comprised 755 species, 8.01% missing values and all traits but milligrams of nectar (~50% of missing values) were included in the imputation process.

@@

We explored the trade-offs between different quantitative plant functional traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson’s correlation = 0.72, _P_ < 0.01 and Pearson’s correlation = 0.36, _P_ < 0.01), and thus we avoided overemphasizing flower size on the reproductive spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th–97.5th percentile range [@legendre2012], and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function _phyl.pca_ from the package _phytools_ version _0.7-70_ [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S2). We found little difference between the explained variance with the imputed dataset (51.08%) and the dataset without missing values (52.87%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

**Phylogenetic signal of traits**

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package _phytools_ [@revell2012] version _0.7-70_ and we used Pagel’s $\lambda$ as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative (presence-absence) and quantitative (interaction frequency) form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with interaction frequency that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and interaction frequency does not consider each pollinator species efficiency [@ballantyne2015], interaction frequency can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version _4.0.3_.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative (presence/absence) and quantitative (visitation rate) floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits of all groups, we also explored the presence-absence of interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that _Apis mellifera_ was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that _A. mellifera_ was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of _A. mellifera_ on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of interaction and visitation rate excluding _A. mellifera_. We found that _A. mellifera_, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package _brms_ [@burkner2017] version _2.14.6_. We modelled the frequency of visits as a function of the main axes of plant trait variation and their interactions with floral visitor functional groups (Visits ~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the floral visitor guilds. In this model, we added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified this model with a zero inflated negative binomial distribution and weakly informative priors from the brms function. We run this model for 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the _bayesplot_ package [@gabry2019] version _1.7.2_.

@@

We investigated whether different quantitative traits determined plant species functional roles using Bayesian modelling and regression trees. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization [d’\; @bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the _specieslevel_ function from the R package _bipartite_ [@dormann2008] version _2.15_.

@@

Second, to better understand these complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on species functional roles. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the _rpart_ package [@therneau2015] version _4.1-15_ with method _‘anova’_ with a minimum of 50 observations per terminal node and we used the _rpart.plot_ package [@milborrow2015] version _3.0.9_ to plot the regression trees. We considered the species level indices as response variables (interaction frequency, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S4).

@@

The phylogenetically informed principal component analysis (pPCA) captured by the first two and three axes 51.8% and 70.97% of trait variation, respectively (Fig. 1 and Fig. S5) and had a phylogenetic correlation ($\lambda$) of 0.76. The first principal component (PC1) represented 26.72% of the trait variation and indicated a trade-off between flower number and flower size. We refer to this axis as the ‘flower number - flower size trade-off’, as already described in previous studies [@sargent2007; @kettle2011]. Hence, one end of the spectrum comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings > |0.5|; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). The second principal component (PC2) represented 25.05% of the trait variation and indicated a trade-off between low and high pollinator dependence. We refer to this axis as the ‘pollinator dependence trade-off’. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). We found that high pollinator dependence was associated with larger and a higher number of flowers, greater plant height and longer styles. In contrast, species with high levels of autonomous selfing tended to have fewer and smaller flowers, had shorter styles and were shorter in height. Further, PC3 explained a considerable amount of trait variability (19.17%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Table S3). Thus, because style length was correlated with all traits on PC3 and was the main driver of trait variation, we refer to this axis as the ‘style length trade-off’. Further, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82%; Fig. S6). This pPCA explained similar variance with the first two principal components (45.52%) and similar associations of traits despite some variability in the loadings (Table S4).

@@

**Phylogenetic signal of traits**

>>>>

\textbf{Number of words:} Abstract  (146 words), main text (4999 words).

\textbf{Number of references:} 75

@@

Plant life-history strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of plant trait variation, little is known about plant reproductive trade-offs and how these constrain life-history strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these drive interactions with floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over half of all plant reproductive trait variation was explained by two independent axes. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. In addition, plant reproductive trade-offs determined important differences in the interaction level among floral visitor guilds. Our study shows the main reproductive trade-offs of flowering plants and their relevance to understand plant-pollinator interactions in a global context.

@@

Despite the astonishing diversity of floral structures among flowering plants [@barrett2002; @schiestl2013] and their importance in shaping plant-pollinator interactions [@fenster2004; @dellinger2020], a unified framework that **describes the major** plant reproductive trade-offs is currently lacking [@roddy2021]. In addition, macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with pollinators at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. **However, plant reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020] and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations [@tur2013; @devaux2014].**

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases [e.g., TRY @kattge2011; and COMPADRE @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies have focused on vegetative traits such as leaf [@wright2004], wood [@chave2009], or root [@laughlin2021] trade-offs with little or no attention given to reproductive traits [@evojtko2020; @roddy2021] which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, [@barrett2003; @moeller2017] and outcrossing levels are positively correlated with flower size [@goodwillie2010]. In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species’ reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators [@ollerton2011; @rodger2021]. However, it is still unknown to what extent these different reproductive compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties [@bartomeus2013; @olito2015; @rowe2020]. Moreover, plant traits can define species’ network roles [e.g., specialists vs generalists\; @lazaro2013; @tur2013]. For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. **Importantly, the combination of traits has shown to increase the predictive power of the network interactions [@eklof2013] and by considering the position in the multidimensional trait space we can determine species’ functional roles within plant-pollinator networks [@dehling2016]. Nonetheless, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.**

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (https://www.ncbi.nlm.nih.gov/taxonomy) for plants and ITIS (https://www.itis.gov/) for pollinators, using the R package _taxize_ [@chamberlain2020]. We filled the ‘not found’ searches manually using http://www.theplantlist.org/ and http://www.mobot.org/ for plants and http://www.catalogueoflife.org/ for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (see Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8% of the dataset, see Fig. S1 for missing values information for each trait).

@@

n_of_records <- subset(missing.values, isna=="FALSE")
Trait <- c("Plant height (m)", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovules per flower",
           "Flowers per plant", "Nectar ($\\mu$l)", "Nectar (mg)", "Nectar concentration ($\\%$)", "Pollen grains per flower", "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Nectar", "Compatibility system", "Breeding system", " ", " ", " ", " ", " ")

T.cat <- linebreak(c("Short-lived \n Perennial", "Herb \n Shrub \n Tree", "Brush \n Campanulate \n Capitulum \n Open \n Papilionaceous \n Tube",  "Actinomorphic \n Zygomorphic",
           "Presence \n Absence", "Self-incomp. \n Part. self-comp. \n Self-comp.", "Hermaphrodite \n Monoecious \n Dioecious"," ", " ", " ", " ", ""),align = c("l"))

@@

Type_1 <- linebreak(c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Reproductive \n biology", "Reproductive \n biology","","","","", ""),align = c("c"))

@@

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE, align = c("c","l","c","c","l","l","c"),caption = "\\textbf{Table 1 | Quantitative and categorical traits used in this study.}") %>%
kable_styling(full_width = F,font_size = 10) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 3),bold=T)  %>%
    column_spec(c(1,3), bold = T) %>% row_spec(0, bold=T)

@@

Trait missing values were imputed with the function _missForest_ [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of _missForest_ [@penone2014]. To extract the eigenvectors, we used the function _PVRdecomp_ from the package _PVR_ [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species, 5.79% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S1) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01% of missing values).

@@

We explored the trade-offs between different quantitative plant functional traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson’s correlation = 0.72, _P_ < 0.01 and Pearson’s correlation = 0.36, _P_ < 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th–97.5th percentile range [@legendre2012], and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function _phyl.pca_ from the package _phytools_ [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (see Fig. S2). We found little difference between the explained variance with the imputed dataset (51.08%) and the dataset without missing values (52.87%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

**Phylogenetic signal**

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package _phytools_ [@revell2012] and we used Pagel’s $\lambda$ as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with interaction frequency that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and interaction frequency does not consider each pollinator species efficiency [@ballantyne2015], interaction frequency can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version _4.0.3_.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored the presence-absence of interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that _Apis mellifera_ was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that _A. mellifera_ was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of _A. mellifera_ on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of interaction and visitation rate excluding _A. mellifera_. We found that _A. mellifera_, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package _brms_ [@burkner2017]. We modelled the frequency of visits as a function of the main axes of plant trait variation and their interactions with floral visitor functional groups (Visits ~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the floral visitor guilds. In this model, we added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified this model with a zero inflated negative binomial distribution and weakly informative priors from the brms function. We run this model for 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the _bayesplot_ package [@gabry2019].

@@

We investigated whether different quantitative traits determined plant species functional roles using Bayesian modelling and regression trees. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization (d’) [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the _specieslevel_ function from the R package _bipartite_ [@dormann2008].

@@

Second, to better understand these complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on species functional roles. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the _rpart_ package [@therneau2015] with method _‘anova’_ with a minimum of 50 observations per terminal node and we used the _rpart.plot_ package [@milborrow2015] to plot the regression trees. We considered the species level indices as response variables (interaction frequency, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S4).

@@

The phylogenetically informed principal component analysis captured by the first two and three axes 51.8% and 70.97% of trait variation, respectively (Fig. 1 and Fig. S5) and had a phylogenetic correlation ($\lambda$) of 0.76. The first principal component (PC1) represented 26.72% of the trait variation and indicated a trade-off between flower number and flower size. We refer to this axis as the ‘flower number - flower size trade-off’, as already described in previous studies [@sargent2007; @kettle2011]. Hence, one end of the spectrum comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings > |0.5|; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). The second principal component (PC2) represented 25.05% of the trait variation and indicated a trade-off between low and high pollinator dependence. We refer to this axis as the ‘pollinator dependence trade-off’. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). We found that high pollinator dependence was associated with larger and a higher number of flowers, greater plant height and longer styles. In contrast, species with high levels of autonomous selfing tended to have fewer and smaller flowers, had shorter styles and were shorter in height. Further, PC3 explained a considerable amount of trait variability (19.17%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Table S3). Thus, because style length was correlated with all traits on PC3 and was the main driver of trait variation, we refer to this axis as the ‘style length trade-off’. Further, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82%; Fig. S6). This pPCA explained similar variance with the first two principal components (45.52%) and similar associations of traits despite some variability in the loadings (Table S4).

@@

**Phylogenetic signal**
